---
output: html_document
editor_options: 
  chunk_output_type: console
---
style>
div.blue pre { background-color:lightblue; font-size:11pt}
div.blue pre.r { background-color:cornsilk; font-size: 11pt}
</style>
<div class = "blue">

---
title: "Capstone Project"
author: "Jeffrey Wong"
date: "03/10/2021"
output: html_document
---
<font size="4">

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/ECON691-01")
options(stringsAsFactors = F)
myPalette01 <- rep(RColorBrewer::brewer.pal(12, 'Set1'), times=3)
myPalette02 <- rep(RColorBrewer::brewer.pal(11, 'PuOr'), times=2)
library(tidyverse)
library(grf)
```


```{r data prep}

data <- read.csv('discon_full_mkt02.csv', row.names = 1, stringsAsFactors = T)
# rearrange the data in chronological order to prevent data leakage
data$Announce.Date <-  as.Date(data$Announce.Date, format='%Y-%m-%d')
data <- data[order(data$Announce.Date, data$acqID, data$day),]
# log the prices
data$log_price <- log(data$price)
data$log_DJ <- log(data$DJ)
data$log_SP500 <- log(data$SP500)
data$log_NASDAQ <- log(data$NASDAQ)
# factorise acq ID
data$acqID <- as.factor(data$acqID)

data <- data[,c(1,6,7,13:ncol(data))]


# split the data into splitting and estimating set
id <- unique(data$acqID)
cases <- id[1:round(length(id) * .6)]
train <- data[which(data$acqID %in% cases), c(1:14)]
test <- data[-which(data$acqID %in% cases), c(1:14)]

cf <- causal_forest(
  X = model.matrix(~ ., data = train[, -c(4,11)]),
  Y = train$log_price,
  W = train$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10)

#predict using the test set
preds <- predict(
  object = cf, 
  newdata = model.matrix(~ ., data = test[, -c(4,11)]), 
  estimate.variance = TRUE)

test$preds <- preds$predictions



ATE <- average_treatment_effect(cf)
paste("95% CI for the ATE:", round(ATE[1], 3), "+/-", round(qnorm(0.975) * ATE[2], 3))


ggplot(test, aes(x=preds)) + geom_histogram() + theme_light() + 
  labs(x='Estimated Treatment', y='Frequency')

high_effect <- (test$preds > median(test$preds))


test_calibration(cf)

features <- cf %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(cf$X.orig)) %>% 
  arrange(desc(V1))

features <- features[features$V1>0.004,]

ggplot(data=features, aes(x=V1, y=reorder(variable, V1))) + geom_point() + 
  theme_light() + labs(x="Importance", y="")


p1 <- ggplot(test, aes(x = log_NASDAQ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()

p2 <- ggplot(test, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()

p3 <- ggplot(test, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

p4 <- ggplot(test, aes(x = libor_1M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

cowplot::plot_grid(p1, p2, p3, p4)

plot_htes(preds, ci=T)
```


```{r quantile analysis}

# split into 4 quantiles depending on the log price before M&A
tiles <- quantile(data[data$M_A==0,]$log_price)

#set the data into 4 quantiles
data$tile <- unlist(lapply(seq(1:dim(data)[1]), function(x)
{
  if (data$log_price[x]<=tiles[[2]])
  {1}
  else if (data$log_price[x]>tiles[[2]] & data$log_price[x]<=tiles[[3]])
  {2}
  else if (data$log_price[x]>tiles[[3]] & data$log_price[x]<=tiles[[4]])
  {3}
  else {4}
}
))
# split data into 4 quantiles
data1 <- data[data$tile==1,]
data2 <- data[data$tile==2,]
data3 <- data[data$tile==3,]
data4 <- data[data$tile==4,]



# split the dataset into train and test
id <- unique(data1$acqID)
cases <- id[1:round(length(id) * .6)]
train1 <- data1[which(data1$acqID %in% cases), c(1:14)]
test1 <- data1[-which(data1$acqID %in% cases), c(1:14)]

id <- unique(data2$acqID)
cases <- id[1:round(length(id) * .6)]
train2 <- data2[which(data2$acqID %in% cases), c(1:14)]
test2 <- data2[-which(data2$acqID %in% cases), c(1:14)]

id <- unique(data3$acqID)
cases <- id[1:round(length(id) * .6)]
train3 <- data3[which(data3$acqID %in% cases), c(1:14)]
test3 <- data3[-which(data3$acqID %in% cases), c(1:14)]

id <- unique(data4$acqID)
cases <- id[1:round(length(id) * .6)]
train4 <- data4[which(data4$acqID %in% cases), c(1:14)]
test4 <- data4[-which(data4$acqID %in% cases), c(1:14)]



```


```{r HTE function}
plot_htes <- function(cf_preds, ci = FALSE, z = 1.96) {
  if (is.null(cf_preds$predictions) )
    stop("cf_preds must include a matrix called 'predictions'")
  
  out <- ggplot(
    mapping = aes(
      x = rank(cf_preds$predictions), 
      y = cf_preds$predictions
    )
  ) +
    geom_point() +
    labs(x = "Rank", y = "Estimated Treatment Effect") +
    theme_light()
  
  if (ci) {
    out <- out +
      geom_errorbar(
        mapping = aes(
          ymin = cf_preds$predictions + z * sqrt(cf_preds$variance.estimates),
          ymax = cf_preds$predictions - z * sqrt(cf_preds$variance.estimates)
        )
      )
  }
    
  return(out)
}
```

#train the causal forest with train set
cf1 <- causal_forest(
  X = model.matrix(~ ., data = train1[, -c(4, 15, 5:8, 19)]),
  Y = train1$log_price,
  W = train1$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10
)



#predict using the test set
preds1 <- predict(
  object = cf1, 
  newdata = model.matrix(~ ., data = test1[, -c(4, 15, 5:8, 19)]), 
  estimate.variance = TRUE
)

test1$preds <- preds1$predictions

plot(test1$log_price, preds1$predictions, xlab = "Log_price", ylab = "preds", type = "l")

ggplot(test1, aes(x = log_price, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()


#check feature importances

features1 <- cf1 %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(cf1$X.orig)) %>% 
  arrange(desc(V1))


features1 <- features1[features1$V1>0.0005,]

ggplot(data=features1, aes(x=V1, y=reorder(variable, V1))) + geom_point() + 
  theme_light()


p1 <- ggplot(test1, aes(x = libor_3M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

p2 <- ggplot(test1, aes(x = log_SP500, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

p3 <- ggplot(test1, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

p4 <- ggplot(test1, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 1) +
  theme_light()

cowplot::plot_grid(p1, p2, p3, p4)



```{r tile 1}

cf1 <- causal_forest(
  X = model.matrix(~ ., data = train1[, -c(4, 11)]),
  Y = train1$log_price,
  W = train1$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10
)

#predict using the test set
preds1 <- predict(
  object = cf1, 
  newdata = model.matrix(~ ., data = test1[, -c(4, 11)]), 
  estimate.variance = TRUE)
test1$preds <- preds1$predictions

ggplot(test1, aes(x=preds)) + geom_histogram() + theme_light() + 
  labs(x='Estimated Treatment', y='Frequency')

test_calibration(cf1)
```



```{r tile 2}
#tile 2

cf2 <- causal_forest(
  X = model.matrix(~ ., data = train2[, -c(4, 11)]),
  Y = train2$log_price,
  W = train2$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10
)



#predict using the test set
preds2 <- predict(
  object = cf2, 
  newdata = model.matrix(~ ., data = test2[, -c(4, 11)]), 
  estimate.variance = TRUE
)

test2$preds <- preds2$predictions

ggplot(test2, aes(x=preds)) + geom_histogram() + theme_light() + 
  labs(x='Estimated Treatment', y='Frequency')

test_calibration(cf2)

```


```{r tile 3}

cf3 <- causal_forest(
  X = model.matrix(~ ., data = train3[, -c(4, 11)]),
  Y = train3$log_price,
  W = train3$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10
)

#predict using the test set
preds3 <- predict(
  object = cf3, 
  newdata = model.matrix(~ ., data = test3[, -c(4, 11)]), 
  estimate.variance = TRUE
)

test3$preds <- preds3$predictions

ggplot(test3, aes(x=preds)) + geom_histogram() + theme_light() + 
  labs(x='Estimated Treatment', y='Frequency')

test_calibration(cf3)

```


```{r tile 4}

cf4 <- causal_forest(
  X = model.matrix(~ ., data = train4[, -c(4, 11)]),
  Y = train4$log_price,
  W = train4$M_A,
  num.trees = 5000,
  seed = 12345,
  min.node.size = 10
)

#predict using the test set
preds4 <- predict(
  object = cf4, 
  newdata = model.matrix(~ ., data = test4[, -c(4, 11)]), 
  estimate.variance = TRUE
)
test4$preds <- preds4$predictions

ggplot(test4, aes(x=preds)) + geom_histogram() + theme_light() + 
  labs(x='Estimated Treatment', y='Frequency')

test_calibration(cf4)
```


```{r treatment effects}
p1 <- ggplot(test1, aes(x = log_NASDAQ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "1st Quantile")

p2 <- ggplot(test1, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "1st Quantile")

p3 <- ggplot(test2, aes(x = log_NASDAQ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "2nd Quantile")

p4 <- ggplot(test2, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "2nd Quantile")

p5 <- ggplot(test3, aes(x = log_NASDAQ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "3rd Quantile")

p6 <- ggplot(test3, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "3rd Quantile")

p7 <- ggplot(test4, aes(x = log_NASDAQ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "4th Quantile")

p8 <- ggplot(test4, aes(x = log_DJ, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "4th Quantile")

cowplot::plot_grid(p1, p2, NULL,
                   p3, p4, NULL,
                   p5, p6, NULL,
                   p7, p8, ncol = 3)


# HTE check
p1 <- ggplot(test1, aes(x = libor_1M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "1st Quantile")

p2 <- ggplot(test1, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "1st Quantile")

p3 <- ggplot(test2, aes(x = libor_1M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light() + labs(title = "2nd Quantile")

p4 <- ggplot(test2, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "2nd Quantile")

p5 <- ggplot(test3, aes(x = libor_1M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "3rd Quantile")

p6 <- ggplot(test3, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "3rd Quantile")

p7 <- ggplot(test4, aes(x = libor_1M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "4th Quantile")

p8 <- ggplot(test4, aes(x = libor_12M, y = preds)) +
  geom_point(size=1) +
  geom_smooth(method = "loess", span = 0.5) +
  theme_light()  + labs(title = "4th Quantile")

cowplot::plot_grid(p1, p2, NULL,
                   p3, p4, NULL,
                   p5, p6, NULL,
                   p7, p8, ncol=3)

```

```{r}
t1 <- plot_htes(preds1, ci=T)
t2 <- plot_htes(preds2, ci=T)
t3 <- plot_htes(preds3, ci=T)
t4 <- plot_htes(preds4, ci=T)

cowplot::plot_grid(t1, t2, t3, t4, labels = c('1st Quantile', 
                                              '2nd Quantile',
                                              '3rd Quantile',
                                              '4th Quantile'))


```

```{r OLS}

variables <- data3[,c(1, 6:8, 10:ncol(data3))]
X <- colnames(variables[, -c(4, 15, 5:8, 19)])

colnames(variables)

OLS <- lm(log_price ~ day + libor_1M + libor_2M + libor_3M + libor_6M +
            libor_12M + log_DJ+log_SP500+log_NASDAQ + M_A, data= data4)
summary(OLS)

```

</font>
